Automated Formal Synthesis\\ of Safe Digital Controllers for Continuous Plants

### 2018-05-16 automatica/paper.tex
fixes

\emph{The first approach} uses a na\"ive technique that starts by devising a
digital controller that stabilizes the plant model while remaining safe for a
pre-selected time horizon and a single initial state; then, it verifies
unbounded-time safety by unfolding the model dynamics, considering the 
full set of initial states, and checking a \emph{completeness
threshold}~\cite{DBLP:conf/vmcai/KroeningS03}, i.e., the number of
iterations required to sufficiently unwind the closed-loop state-space
model such that the safety boundaries are not violated for any larger number of
iterations.  Requiring unfolding up to the completeness threshold, this
approach is computationally expensive.

@@

%%   \note{[this statement is way too broad] Existing methods for controller synthesis rely on transfer 
%%   function representations, which are inadequate to prove safety requirements. } 
%% %
\item We provide two novel algorithms: the first, a na\"ive approach, relies on 
  an unfolding of the dynamics up to a completeness threshold, while the
  second one is abstraction-based and leverages abstraction/refinement and
  acceleration to improve scalability while retaining soundness.  %% A third
  %% algorithm evaluates systems using transfer functions where no state-space
  %% description is given.
  Both approaches provide sound synthesis of 
  feedback controllers and consider the various sources of imprecision in
  the implementation of the control algorithm and in the modeling of the
  plant. 
%
\item We develop a model for different sources of quantization error and account for 
  their effect on safety.  We give bounds that ensure the
  safety of our controllers in a hybrid (i.e., continuous/digital) domain.

@@

We recall a technique called \emph{Abstract Acceleration}~\cite{JSS14,cattaruzza2015unbounded}, 
which will be key in this work for abstraction-based synthesis.  
Abstract Acceleration allows one to perform reachability analysis of dynamical models, 
and as such can be employed for safety verification. 
% 
Given an autonomous model of an iterative program with dynamics 
%
$\vec{x}_{k+1}=\mat{A_d}\vec{x}_k$, 
where $\vec{x}_k \in \mathbb{R}^p$, $\mat{A_d} \in \mathbb{R}^{p \times p}$ and $k \in \mathbb N_0$, 
%\end{equation}
%
we want to find all possible states visited over an infinite time horizon starting at a given initial set $X_0$.  
This set is known as the \emph{reach tube} of the model: 
%
\begin{equation}
\hat{X} = \{ \vec{x}_k: k \in \mathbb N_0, \vec{x}_0 \in X_0 \text{ and } \vec{x}_{k+1}=\mat{A_d}\vec{x}_k\}.
\end{equation}
%
Using Abstract Acceleration we transform this sequential reachability problem into a one-step problem. 
This has a lot of benefits ~\cite{JSS14}, not least in terms of numerical error propagation.   
In order to attain this, 
the problem statement is modified 
%, by means of acceleration, 
into 
%
\begin{equation}
\hat{X} = \{ \vec{x}_k=\mat{A_d}^k\vec{x}_0: k \in N_0 \text{ and } \vec{x}_0 \in X_0 \}, 
\end{equation}
%
and a new 
%model semantics and corresponding 
set $\hat{X}^\sharp$ is further defined as 
%
\begin{equation}\label{eq:aa_reachtube}
\hat{X} \subseteq \hat{X}^\sharp = \mathcal{A}X_0, \text{ such that } \bigcup_{k \in [0\ \infty)} \mat{A_d}^k \subseteq \mathcal{A}, 
\end{equation}
%
where the newly introduced \emph{abstract reach tube} $\hat{X}^\sharp$ is an
over-approximation of the reach tube of the model: 
\cite{JSS14} details algorithms to compute such set, 
whereas \cite{cattaruzza2015unbounded} extends this to LTI models with inputs as in \note{Eqn. XX}.  
In view of the overall goal of this work, 
we extend the algorithm to continuous-time models.  
% and  For this reason we must first look at the acceleration of continuous-time dynamics.

@@

We next describe how to obtain an over-approximation of such a set. 
This is defined as a polyhedron, expressed as an intersection of hyperplanes. 
A tight polyhedron that can be obtained with faces (hyperplanes) that touch the set. 
\reply{This is achieved by picking a number of points in the set, and finding a hyperplane
tangent to that point w.r.t. dynamics.  Inorder to simplify this procedure, we
look at two-dimensional cuts of the space, one at a time (i.e., for every pair
of eigenvalues of the closed loop dynamics, we find a hyperplane tangent to
the progression of that pair in a two dimensional space and use it as a face of
the polyhedron).  The difference with~\cite{cattaruzza2015unbounded} is that
in the continuous case, the tangent is found by derivation rather than
differentialtion, thus including all intermediate points in the arc.}

@@

In this section we discuss two instantiations of CEGIS: 
the first one,
named {\it multi-stage} approach, 
relies on an unfolding of the dynamics up to a completeness threshold; 
the second one is {\it abstraction-based}, 
and it leverages abstraction-refinement \note{ cit.} and acceleration \note{ref to relevant section} to improve scalability,  
while retaining soundness.

@@

To synthesize a controller that satisfies the
stability criteria, we require that a computed polynomial satisfies Jury's
criterion~\cite{fadali} (\reply{see Section~\ref{ssec:stabspecification}}).

Regarding the second requirement, 
we fix an index $k$, 
and we synthesize a safe controller by
unfolding the transition system $k$ steps and by selecting a controller $K$ and a single initial state, 
such that the states at each step do not violate the safety criteria (\reply{see Section~\ref{ssec:safespecification}}).  
That is, we ask the bounded model checker \note{have we discussed the BMC earlier?} if there exists a $K$ that is safe for at least one $x_0$ in our set of all possible initial states.  
This is sound if the current $k$ is greater than the completeness threshold (see later step).  
We~also assume some (fixed or floating point) precision $\langle I_p,F_p\rangle$ for the plant and a sampling rate.  
The checks that these assumptions hold are performed by subsequent {\sc verify} stages. 
\note{we have not discussed the bounds on the input signal. }

@@

%  The na\"ive approach described in Sec.~\ref{sec:CEGIS-precision-incrementation}
  %uses a multi-staged verification phase and performs solution generalization based on
%the plant precision, the number of loop unrollings, 
  %and the sampling rate. Essentially, it
%  synthesizes a controller for an individual initial state
%and input with a bounded time horizon and, subsequently, it generalizes it to all reachable states,
%inputs, and time horizons during the verification phase.

@@

%% Our na\"ive approach
%% creates an abstraction of the system by establishing bounds on the
%% error between such a system and a discrete-time/continuous-space
%% model.  This model is still however, quite complex, hard to verify,
%% and many of the states involved are inconsequential to the safety
%% specification.

@@

The CEGIS loop for this approach is illustrated in Figure~\ref{fig:CEGARIS}. 
%
% Now that we have explained the verification process, we may proceed to describe the synthesis procedure, which is illustrated in figure \ref{fig:CEGARIS}.
% Our model, rather than look at a bisimulation \todo{specific notion that has not been formally define - do not use} in discrete time, explores a much higher abstraction, 
% evaluating the unbounded time-continuous space reach tube as a set of inequalities that represent only the safety specification.
% We have also identified 
% the quantization noise as a bounded non-deterministic input.
It comprises the following steps:  
\begin{enumerate}
\item 
  We start by doing some preprocessing:%%  that 
  %% can be computationally intensive and we don't want it to be part of the CEGIS loop.
% Let us recall the formulas used for single-input and single-output (SISO) systems 
  % in sections \ref{sec:reachable} and \ref{sec:observable}.
  \begin{enumerate}
\item Compute the characteristic polynomial of the matrix $(A_d-B_dK)$ as 
$P_a(z) = z^n+\sum_{i=1}^n{(a_i-k_i)z^{n-i}}$.  As with the na\"ive approach, this polynomial must satisfy Jury's criterion.
\note{are we assuming to be working with a controllable form? should we emphasize that some elements are known, others are parameters? also, why do we use a different notation for this char poly than before?
\reply{[we don't need to talk about controllable form a we only use the characteristic polynomial, but we describe it in more detail because we work directly with the shape indicated here (a-k).}}   
%\blue{[shouldn't there be a b term in the polynomial's coefficients here? No. the polynomial only evaluates the poles]}
%%   A controllable system with quantization will have the following dynamics, where
%% we use underscript ``n'' to denote a noise term,
%% ``cf'' to denote Controllable Canonical Form,
%% $\nu_1$, $\nu_2$ are quantization errors
%% caused by the ADC and DAC conversions, respectively, 
%% and $\nu_3$ is the round-off error.
%% \todo{Why do we use $_t$ here?}
%% %
%% \begin{align}
%% \label{eq:observer_LTI_cf}
%% x_{k+1}=&\mat{A}_{t}x_k+\mat{B}_{t} r_k+\mat{B}_{n} (\nu_1 + \nu_2 + \nu_3)
%% \end{align}
%% \begin{align*}
%% \mat{A}_{t}=&\mat{T}\mat{A}_d\mat{T}^{-1}&
%% \mat{B}_{t}=&\mat{T}\mat{B}_d&
%% \mat{B}_{n}=&[1 \cdots 1]^T&
%% \mat{T}=&\mat{W}_{cf}\mat{W}^{-1}&
%% \end{align*}

@@

In this paper we will use $\mathcal{F}_{\langle I_c,F_c \rangle}(x)$ to denote a real
number $x$ represented at the fixed-point precision of the controller, and
$\mathcal{F}_{\langle I_p,F_p \rangle}(x)$ to denote a real number $x$
represented at the fixed-point precision of the plant model ($I_c$ and $F_c$
are determined by the controller.  We pick $I_p$ and $F_p$ for our synthesis
such that $I_p \geq I_c$ and $\allowbreak F_p \geq F_c$, \reply{so that our controller can be represented accurately in the plant domain}).  
Thus any mathematical operations in our modelled digital controller will be in the
range of $\mathcal{F}_{\langle I_c,F_c \rangle}$, and all other calculations
in our model will be carried out in the range of $\mathcal{F}_{\langle
I_p,F_p \rangle}$.  The physical plant operates in the reals, which means
our verification phase must also account for the numerical error and
quantization errors caused by representing the physical plant at the finite
precision $\mathcal{F}_{\langle I_p,F_p \rangle}$.

>>>>

\emph{The first approach} uses a multi-staged technique that starts by devising a
digital controller that stabilizes the plant model while remaining safe for a
pre-selected time horizon and a single initial state; then, it verifies
unbounded-time safety by unfolding the model dynamics, considering the 
full set of initial states, and checking a \emph{completeness
threshold}~\cite{DBLP:conf/vmcai/KroeningS03}, i.e., the number of
iterations required to sufficiently unwind the closed-loop state-space
model such that the safety boundaries are not violated for any larger number of
iterations.  Requiring unfolding up to the completeness threshold, this
approach is computationally expensive.

@@

%%   \note{[this statement is way too broad] Existing methods for controller synthesis rely on transfer 
%%   function representations, which are inadequate to prove safety requirements. } 
%% %
\item We provide two novel algorithms: the first, a multi-staged approach, relies on 
  an unfolding of the dynamics up to a completeness threshold, while the
  second one is abstraction-based and leverages abstraction/refinement and
  acceleration to improve scalability while retaining soundness.  %% A third
  %% algorithm evaluates systems using transfer functions where no state-space
  %% description is given.
  Both approaches provide sound synthesis of 
  feedback controllers and consider the various sources of imprecision in
  the implementation of the control algorithm and in the modeling of the
  plant. 
%
\item We develop a model for different sources of quantization error and account for 
  their effect on safety.  We give bounds that ensure the
  safety of our controllers in a hybrid (i.e., continuous/digital) domain.

@@

We recall a technique called \emph{Abstract Acceleration}~\cite{JSS14,cattaruzza2015unbounded}, 
which will be key in this work for abstraction-based synthesis.  
Abstract Acceleration allows one to perform reachability analysis of dynamical models, 
and as such can be employed for safety verification. 
% 
Given an autonomous model of an iterative program with dynamics 
%
$\vec{x}_{k+1}=\mat{A_d}\vec{x}_k$, 
where $\vec{x}_k \in \mathbb{R}^p$, $\mat{A_d} \in \mathbb{R}^{p \times p}$ and $k \in \mathbb N_0$, 
%\end{equation}
%
we want to find all possible states visited over an infinite time horizon starting at a given initial set $X_0$.  
This set is known as the \emph{reach tube} of the model: 
%
\begin{equation}
\hat{X} = \{ \vec{x}_k: k \in \mathbb N_0, \vec{x}_0 \in X_0 \text{ and } \vec{x}_{k+1}=\mat{A_d}\vec{x}_k\}.
\end{equation}
%
Using Abstract Acceleration we transform this sequential reachability problem into a one-step problem. 
This has a lot of benefits ~\cite{JSS14}, not least in terms of numerical error propagation.   
In order to attain this, 
the problem statement is modified 
%, by means of acceleration, 
into 
%
\begin{equation}
\hat{X} = \{ \vec{x}_k=\mat{A_d}^k\vec{x}_0: k \in N_0 \text{ and } \vec{x}_0 \in X_0 \}, 
\end{equation}
%
and a new 
%model semantics and corresponding 
set $\hat{X}^\sharp$ is further defined as 
%
\begin{equation}\label{eq:aa_reachtube}
\hat{X} \subseteq \hat{X}^\sharp = \mathcal{A}X_0, \text{ such that } \bigcup_{k \in [0\ \infty)} \mat{A_d}^k \subseteq \mathcal{A}, 
\end{equation}
%
where the newly introduced \emph{abstract reach tube} $\hat{X}^\sharp$ is an
over-approximation of the reach tube of the model: 
\cite{JSS14} details algorithms to compute such set, 
whereas \cite{cattaruzza2015unbounded} extends this to LTI models with inputs as in Eqn. \eqref{eq:plant}.  
In view of the overall goal of this work, 
we extend the algorithm to continuous-time models.  
% and  For this reason we must first look at the acceleration of continuous-time dynamics.

@@

We next describe how to obtain an over-approximation of such a set. 
This is defined as a polyhedron, expressed as an intersection of hyperplanes. 
A tight polyhedron can be obtained with hyperplanes that are tangent to the set,  
which is achieved by selecting a number of points \note{on the boundary} of the set, 
and finding corresponding hyperplanes that are tangent to those points with respect to the dynamics.  
\note{[you are essentially taking the Lie derivative of the set wrt the dynamics at the selected point?]}
In order to simplify this procedure, we look at two-dimensional cuts \note{projection?} of the space, one at a time (i.e., for every pair
of eigenvalues of the closed-loop dynamics, we find a hyperplane tangent to
the progression of that pair in a two dimensional space and use it as a face of
the polyhedron).  The difference with~\cite{cattaruzza2015unbounded} is that
in the continuous case, the tangent is found by derivation \note{cf. `Lie derivative' above} rather than
differentialtion, thus including all intermediate points \note{[?] in the arc}.

@@

In this section we discuss two instantiations of CEGIS: 
the first one,
named {\it multi-stage} approach, 
relies on an unfolding of the dynamics up to a completeness threshold; 
the second one is {\it abstraction-based}, 
and it leverages abstraction-refinement \cite{DBLP:conf/cav/ClarkeGJLV00} and acceleration (cf. Sec.~\ref{ssec:aa}) to improve scalability,  
while retaining soundness.

@@

To synthesize a controller that satisfies the
stability criteria, we require that a computed polynomial satisfies Jury's
criterion~\cite{fadali} (see Section~\ref{ssec:stabspecification}).

Regarding the second requirement, 
we fix an index $k$, 
and we synthesize a safe controller by
unfolding the transition system $k$ steps and by selecting a controller $K$ and a single initial state, 
such that the states at each step do not violate the safety criteria (see Section~\ref{ssec:safespecification}).    
That is, we ask the bounded model checker \note{have we discussed the BMC earlier?} if there exists a $K$ that is safe for at least one $x_0$ in our set of all possible initial states.  
This is sound if the current $k$ is greater than the completeness threshold (see later step).  
We~also assume some (fixed or floating point) precision $\langle I_p,F_p\rangle$ for the plant and a sampling rate.  
The checks that these assumptions hold are performed by subsequent {\sc verify} stages. 
\note{we have not discussed the bounds on the input signal. }

@@

%  The multi-staged approach described in Sec.~\ref{sec:CEGIS-precision-incrementation}
  %uses a multi-staged verification phase and performs solution generalization based on
%the plant precision, the number of loop unrollings, 
  %and the sampling rate. Essentially, it
%  synthesizes a controller for an individual initial state
%and input with a bounded time horizon and, subsequently, it generalizes it to all reachable states,
%inputs, and time horizons during the verification phase.

@@

%% Our multi-staged approach
%% creates an abstraction of the system by establishing bounds on the
%% error between such a system and a discrete-time/continuous-space
%% model.  This model is still however, quite complex, hard to verify,
%% and many of the states involved are inconsequential to the safety
%% specification.

@@

The CEGIS loop for this approach is illustrated in Figure~\ref{fig:CEGARIS}. 
%
% Now that we have explained the verification process, we may proceed to describe the synthesis procedure, which is illustrated in figure \ref{fig:CEGARIS}.
% Our model, rather than look at a bisimulation \todo{specific notion that has not been formally define - do not use} in discrete time, explores a much higher abstraction, 
% evaluating the unbounded time-continuous space reach tube as a set of inequalities that represent only the safety specification.
% We have also identified 
% the quantization noise as a bounded non-deterministic input.
It comprises the following steps:  
\begin{enumerate}
\item 
  We start by doing some preprocessing:%%  that 
  %% can be computationally intensive and we don't want it to be part of the CEGIS loop.
% Let us recall the formulas used for single-input and single-output (SISO) systems 
  % in sections \ref{sec:reachable} and \ref{sec:observable}.
  \begin{enumerate}
\item Compute the characteristic polynomial of the matrix $(A_d-B_dK)$ as 
$P_a(z) = z^n+\sum_{i=1}^n{(a_i-k_i)z^{n-i}}$.  As with the multi-staged approach, this polynomial must satisfy Jury's criterion.
\note{are we assuming to be working with a controllable form? should we emphasize that some elements are known, others are parameters? also, why do we use a different notation for this char poly than before?
\reply{[we don't need to talk about controllable form a we only use the characteristic polynomial, but we describe it in more detail because we work directly with the shape indicated here (a-k).}}   
\note{but, this very shape of the characteristic polynomial holds only for LTI models in controllable canonical form.}
%\blue{[shouldn't there be a b term in the polynomial's coefficients here? No. the polynomial only evaluates the poles]}
%%   A controllable system with quantization will have the following dynamics, where
%% we use underscript ``n'' to denote a noise term,
%% ``cf'' to denote Controllable Canonical Form,
%% $\nu_1$, $\nu_2$ are quantization errors
%% caused by the ADC and DAC conversions, respectively, 
%% and $\nu_3$ is the round-off error.
%% \todo{Why do we use $_t$ here?}
%% %
%% \begin{align}
%% \label{eq:observer_LTI_cf}
%% x_{k+1}=&\mat{A}_{t}x_k+\mat{B}_{t} r_k+\mat{B}_{n} (\nu_1 + \nu_2 + \nu_3)
%% \end{align}
%% \begin{align*}
%% \mat{A}_{t}=&\mat{T}\mat{A}_d\mat{T}^{-1}&
%% \mat{B}_{t}=&\mat{T}\mat{B}_d&
%% \mat{B}_{n}=&[1 \cdots 1]^T&
%% \mat{T}=&\mat{W}_{cf}\mat{W}^{-1}&
%% \end{align*}

@@

In this paper we will use $\mathcal{F}_{\langle I_c,F_c \rangle}(x)$ to denote a real
number $x$ represented at the fixed-point precision of the controller, and
$\mathcal{F}_{\langle I_p,F_p \rangle}(x)$ to denote a real number $x$
represented at the fixed-point precision of the plant model ($I_c$ and $F_c$
are determined by the controller.  We pick $I_p$ and $F_p$ for our synthesis
such that $I_p \geq I_c$ and $\allowbreak F_p \geq F_c$, so that our controller can be represented accurately in the plant domain).  
Thus any mathematical operations in our modelled digital controller will be in the
range of $\mathcal{F}_{\langle I_c,F_c \rangle}$, and all other calculations
in our model will be carried out in the range of $\mathcal{F}_{\langle
I_p,F_p \rangle}$.  The physical plant operates in the reals, which means
our verification phase must also account for the numerical error and
quantization errors caused by representing the physical plant at the finite
precision $\mathcal{F}_{\langle I_p,F_p \rangle}$.
